Mixture Models for Single-Cell Assays with Applications to Vaccine Studies

### 2013-06-25 MIMOSA/mimosafigures/manuscriptFigures.Rmd
Update code to use bootstrap etc..

```{r }
if(!file.exists("hvtnEMfit.rds")){
  hvtn.result.EM<-MIMOSA(NSUB+CYTNUM~ASSAYID+VISITNO+PTID+RX_CODE+RefTreat|TCELLSUB+ANTIGEN+CYTOKINE,hvtndata,ref=RefTreat%in%"Reference"&ANTIGEN%in%"ENV-1-PTEG",subset=RefTreat%in%"Treatment"&ANTIGEN%in%"ENV-1-PTEG",method="EM",run.parallel=TRUE)
  saveRDS(hvtn.result.EM,file="hvtnEMfit.rds")
}else{
  hvtn.result.EM<-readRDS("hvtnEMfit.rds")
}
if(!file.exists("hvtnMCMCfit.rds")){
  hvtn.result.mcmc<-MIMOSA(NSUB+CYTNUM~ASSAYID+VISITNO+PTID+RX_CODE+RefTreat|TCELLSUB+ANTIGEN+CYTOKINE,hvtndata,ref=RefTreat%in%"Reference"&ANTIGEN%in%"ENV-1-PTEG",subset=RefTreat%in%"Treatment"&ANTIGEN%in%"ENV-1-PTEG",method="mcmc",run.parallel=TRUE)
  saveRDS(hvtn.result.mcmc,file="hvtnMCMCfit.rds")
}else{
  hvtn.result.mcmc<-readRDS("hvtnMCMCfit.rds")
}
```

@@

```{r cache=TRUE,fig.width=10,fig.height=8,fig.cap='The TNF expressing cytokine specific cells have high background as evidenced by the poor performance of all methods fit to TNF+ subsets of cells, with the exception of the triple-positive cell subset. As the number of simultaneously expressed cytokines being examined increases, the background decreases.'}
ROC$Method<-relevel(ROC$Method,"MIMOSA (EM)")
ggplot(subset(ROC,TCELLSUB%in%"cd3+/cd4+"))+geom_line(aes(x=FPR,y=TPR,color=Method))+theme_bw()+facet_wrap(~CYTOKINE)+geom_text(aes(x=x,y=y,label=sprintf("AUC=%s",signif(AUC,2)),col=Method),size=3,data=AUC,show_guide=FALSE)+theme(axis.text.x=element_text(angle=90,hjust=1))
```{r fig.width=10,fig.height=8,fig.cap='The FDR is well controlled for cytokine subsets with low background. The TNF expressing subsets have high background and consequently positive assay results in the "Visit 2" samples, which are labelled as "true negatives" in this scheme. The result is observed as a large observed false discovery rate for all methods.'}
FDR$Method<-relevel(FDR$Method,"MIMOSA (EM)")
ggplot(subset(FDR,TCELLSUB%in%c("cd3+/cd4+")))+geom_line(aes(y=true.fdr,x=fdr,color=Method))+geom_abline(1,lty=3)+theme_bw()+scale_x_continuous(name="Nominal FDR")+scale_y_continuous(name="Observed FDR")+facet_wrap(~CYTOKINE)+theme(axis.text.x=element_text(angle=90,hjust=1))
```

@@

```{r heatmap.probs.bygene,cache=TRUE}
foo<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)x@z[,2]))
props<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)do.call(c,lapply(x@result@p,function(x)-diff(x[,2])))))
S<-sign(props)
S[S==0]<-1
colnames(S)<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])[,c("Patient.ID")]
stim<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])$Stim
rownames(S)<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)as.character(unique(pData(x)$Gene))))
heatmap.2((S*foo)[apply(foo,1,function(x)sum(x>.1)>=2),],col=colorpanel(n=10,high="green",low="red",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="signed Posterior Probabilities\n(fit by gene, all stimulations)")
```

@@

```{r heatmap.props.bygene,cache=TRUE}
colnames(props)<-colnames(S)
rownames(props)<-rownames(S)
heatmap.2(S*sqrt(abs(props)),col=colorpanel(n=10,high="green",low="red",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="Posterior Differences in Proportions\n(fit by gene all stimulations)")
```

@@

```{r heatmap.probs.bystim.bygene,cache=TRUE}
foo<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene,function(x)do.call(rbind,lapply(x,function(x)x@z[,2]))))
props<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene,function(x)do.call(rbind,lapply(x,function(x)do.call(c,lapply(x@result@p,function(x)-diff(x[,2])))))))
S<-sign(props)
colnames(S)<-do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])[,c("Patient.ID")])))
S[S==0]<-1
stim<-as.numeric(factor(do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])$Stim)))))
colnames(S)<-do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])$Patient.ID)))
rownames(S)<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene[[1]],function(x)as.character(unique(pData(x)$Gene))))
heatmap.2((S*foo)[apply(foo,1,function(x)sum(x>.1)>=1),],col=colorpanel(n=10,high="green",low="red",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],main="Signed Posterior Probabilities\n(fit by gene and stimulation).")
```

@@

```{r heatmap.props.bystim.bygene,cache=TRUE}
colnames(props)<-colnames(S)
rownames(props)<-rownames(S)
heatmap.2(sqrt(abs(props))*S,col=colorpanel(n=10,high="green",low="red",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="Posterior differences in proportions\n(fit by gene and stimulations)")
```

@@

```{r cache=TRUE}
fisher<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x){
  p<-vector('numeric',nrow(x@result@n.stim))
  for(i in 1:nrow(x@result@n.stim)){
    p[i]<-fisher.test(as.matrix(rbind(x@result@n.stim[i,],x@result@n.unstim[i,])),alternative="two.sided")$p.value
  }
  p
}))
S<-sign(lapply(fluidigm.fits.by.stim.allgenes,function(x)do.call(rbind,lapply(x,function(x)(prop.table(as.matrix(x@result@n.stim),1)-prop.table(as.matrix(x@result@n.unstim),1))[,2])))[[1]])
S[S==0]<-1
colnames(fisher)<-gsub("_B3GAT1_Treatment","",colnames(S))
rownames(fisher)<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)as.character(unique(pData(x)$Gene))))
stim<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])$Stim
heatmap.2(fisher*S,col=colorpanel(n=10,high="green",low="red",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="-log10(p-values)\nfrom Fisher's Exact Test")
```

@@

toplot<-data.frame(rbind(data.frame(CC$n.stim),data.frame(CC$n.unstim)),stim=c(rep("Stim",nrow(CC$n.stim)),rep("Unstim",nrow(CC$n.unstim))))
toplot$id<-rep(unlist(lapply(strsplit(as.character(subset(C,Stim.Condition%in%"Unstim")[,1]),":"),function(x)x[1]),use.names=FALSE),2)
colnames(toplot)[1:4]<-gsub("X","",paste(g1,g2,colnames(toplot)[1:4],sep=":"))
toplot<-melt(toplot,id=c("id","stim"))
ggplot(toplot)+geom_raster(aes(x=variable,y=id,fill=value))+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous(na.value="gray",breaks=round(seq(0,96,l=10)))

DD<-CC
DD[[1]]<-cbind(CC[[1]][,1],rowSums(CC[[1]][,2:4]))
DD[[2]]<-cbind(CC[[2]][,1],rowSums(CC[[2]][,2:4]))
toplot2<-data.frame(rbind(data.frame(DD$n.stim),data.frame(DD$n.unstim)),stim=c(rep("Stim",nrow(DD$n.stim)),rep("Unstim",nrow(DD$n.unstim))))
toplot2$id<-rep(unlist(lapply(strsplit(as.character(subset(C,Stim.Condition%in%"Unstim")[,1]),":"),function(x)x[1]),use.names=FALSE),2)
colnames(toplot2)[1:2]<-gsub("X","",paste(g1,g2,colnames(toplot2)[1:2],sep=":"))
toplot2<-melt(toplot2,id=c("id","stim"))
ggplot(toplot2)+geom_raster(aes(x=variable,y=id,fill=value))+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous(na.value="gray",breaks=round(seq(0,96,l=10)))

```{r fitmulti,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE,results='hide'}
flres<-.fitMCMC(data=CC,inits=MIMOSA:::MDMix(CC,initonly=TRUE),EXPRATE=1e-2,pXi=1,alternative="not equal")
flsum<-.fitMCMC(data=DD,inits=MIMOSA:::MDMix(DD,initonly=TRUE),EXPRATE=1e-2,pXi=1,alternative="not equal")
sapply(list.files(pattern=c("\\.dat")),unlink)

@@

FDR<-function(x,truth){
  o<-order(x,decreasing=FALSE)
  cbind(tpr=cumsum(truth[o])/sum(truth),fpr=cumsum(!truth[o])/sum(!truth))
}

fisher.p.adj<-lapply(fisher.p,function(x)p.adjust(x,"fdr"))
fisher.roc<-do.call(rbind,by(cbind(do.call(rbind,lapply(fisher.p.adj,function(i)FDR(i,gl(2,50)==2)))),rep(gl(100,1),10),function(x)colMeans(x)))
mimosa.adj<-lapply(1:ncol(mdsim),function(i)MIMOSA:::fdr(mdsim[,i]$z))
MIMOSA.roc<-do.call(rbind,by(do.call(rbind,lapply(mimosa.adj,function(i)FDR(i,gl(2,50)==2))),rep(gl(100,1),10),function(x)colMeans(x)))
foo<-data.frame(as.data.frame(rbind(fisher.roc,MIMOSA.roc)),Method=gl(2,100,labels=c("Fisher","MIMOSA")))
sz<-element_text(size=18)
ggplot(foo)+geom_line(aes(x=fpr,y=tpr,col=Method))+scale_y_continuous("TPR")+scale_x_continuous("FPR")+theme_bw()+theme(axis.title.x=sz,axis.title.y=sz,axis.text.x=sz,axis.text.y=sz)
```

@@

>>>>

```{r }
if(!file.exists("hvtnEMfit.rds")){
  hvtn.result.EM<-MIMOSA(NSUB+CYTNUM~ASSAYID+VISITNO+PTID+RX_CODE+RefTreat|TCELLSUB+ANTIGEN+CYTOKINE,hvtndata,ref=RefTreat%in%"Reference"&ANTIGEN%in%"ENV-1-PTEG",subset=RefTreat%in%"Treatment"&ANTIGEN%in%"ENV-1-PTEG",method="EM",run.parallel=TRUE)
  saveRDS(hvtn.result.EM,file="hvtnEMfit.rds")
}else{
  hvtn.result.EM<-readRDS("hvtnEMfit.rds")
}
if(!file.exists("hvtnMCMCfit.rds")){
  hvtn.result.mcmc<-MIMOSA(NSUB+CYTNUM~ASSAYID+VISITNO+PTID+RX_CODE+RefTreat|TCELLSUB+ANTIGEN+CYTOKINE,hvtndata,ref=RefTreat%in%"Reference"&ANTIGEN%in%"ENV-1-PTEG",subset=RefTreat%in%"Treatment"&ANTIGEN%in%"ENV-1-PTEG",method="mcmc",run.parallel=TRUE,pXi=0.1,getP=TRUE)
  saveRDS(hvtn.result.mcmc,file="hvtnMCMCfit.rds")
}else{
  hvtn.result.mcmc<-readRDS("hvtnMCMCfit.rds")
}
```

@@

```{r cache=TRUE,fig.width=10,fig.height=8,fig.cap='The TNF expressing cytokine specific cells have high background as evidenced by the poor performance of all methods fit to TNF+ subsets of cells, with the exception of the triple-positive cell subset. As the number of simultaneously expressed cytokines being examined increases, the background decreases.'}
ROC$Method<-relevel(ROC$Method,"MIMOSA (EM)")
levs <-
  c("IFN-IL2-TNF+", "IFN-IL2+TNF-", "IFN-IL2+TNF+", "IFN+IL2-TNF-", 
    "IFN+IL2-TNF+", "IFN+IL2+TNF-", "IFN+IL2+TNF+", "IFNg+", "IFNg+IL2+", 
    "IFNg+TNF+", "IL2+", "IL2-IFNg+", "IL2+ OR IFNg+", "IL2+IFNg-", 
    "IL2+IFNg+", "IL2+TNF+", "TNF+")
ROC$CYTOKINE<-factor(ROC$CYTOKINE,labels=levs)
AUC$CYTOKINE<-factor(AUC$CYTOKINE,labels=levs)
rocplots<-ggplot(subset(ROC,TCELLSUB%in%"cd3+/cd4+"))+geom_line(aes(x=FPR,y=TPR,color=Method))+theme_bw()+facet_wrap(~CYTOKINE)+geom_text(aes(x=x,y=y,label=sprintf("AUC=%s",signif(AUC,2)),col=Method),size=3,data=AUC,show_guide=FALSE)+theme(axis.text.x=element_text(angle=90,hjust=1))
```{r fig.width=10,fig.height=8,fig.cap='The FDR is well controlled for cytokine subsets with low background. The TNF expressing subsets have high background and consequently positive assay results in the "Visit 2" samples, which are labelled as "true negatives" in this scheme. The result is observed as a large observed false discovery rate for all methods.'}
FDR$Method<-relevel(FDR$Method,"MIMOSA (EM)")
levs<-levels(FDR$CYTOKINE)
levs <-
  c("IFN-IL2-TNF+", "IFN-IL2+TNF-", "IFN-IL2+TNF+", "IFN+IL2-TNF-", 
    "IFN+IL2-TNF+", "IFN+IL2+TNF-", "IFN+IL2+TNF+", "IFNg+", "IFNg+IL2+", 
    "IFNg+TNF+", "IL2+", "IL2-IFNg+", "IL2+ OR IFNg+", "IL2+IFNg-", 
    "IL2+IFNg+", "IL2+TNF+", "TNF+")
FDR$CYTOKINE<-factor(FDR$CYTOKINE,labels=levs)
fdrplots<-ggplot(subset(FDR,TCELLSUB%in%c("cd3+/cd4+")))+geom_line(aes(y=true.fdr,x=fdr,color=Method))+geom_abline(1,lty=3)+theme_bw()+scale_x_continuous(name="Nominal FDR")+scale_y_continuous(name="Observed FDR")+facet_wrap(~CYTOKINE)+theme(axis.text.x=element_text(angle=90,hjust=1))
```

Figure 1 and S1. Write to Figures folder.
Don't evaluate here.

```{r echo=FALSE,eval=FALSE}
suppressPackageStartupMessages(library(gridExtra))
f1levs<-levs[c(8,11,13,17,15,16,10,7)]
sflevs<-setdiff(levs,f1levs)
AUC$y[AUC$Method%in%"Fisher"]<-0.4
AUC$y[AUC$Method%in%"MIMOSA (EM)"]<-0.5
AUC$y[AUC$Method%in%"MIMOSA (MCMC)"]<-0.3

fnames<-c("revised_2.pdf","revised_1.pdf","revised_15.pdf","revised_3.pdf","revised_5.pdf","revised_6.pdf","revised_7.pdf","revised_8.pdf")
for(i in seq_along(f1levs)){
  fi.1<-ggplot(subset(ROC,TCELLSUB%in%"cd3+/cd4+"&CYTOKINE%in%f1levs[i]))+geom_line(aes(x=FPR,y=TPR,color=Method))+theme_bw()+facet_wrap(~CYTOKINE)+geom_text(aes(x=x,y=y,label=sprintf("AUC=%s",signif(AUC,2)),col=factor(Method)),data=subset(AUC,CYTOKINE%in%f1levs[i]),show_guide=FALSE)+theme(axis.text.x=element_text(angle=90,hjust=1)) 
  fi.2<-ggplot(subset(FDR,TCELLSUB%in%c("cd3+/cd4+")&CYTOKINE%in%f1levs[i]))+geom_line(aes(y=true.fdr,x=fdr,color=Method))+geom_abline(1,lty=3)+theme_bw()+scale_x_continuous(name="Nominal FDR")+scale_y_continuous(name="Observed FDR")+facet_wrap(~CYTOKINE)+theme(axis.text.x=element_text(angle=90,hjust=1))
  pdf(file=sprintf("/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/%s",fnames[i]),width=10,height=3)
  grid.arrange(fi.1,fi.2,ncol=2,nrow=1)
  dev.off()
  grid.arrange(fi.1,fi.2,ncol=2,nrow=1)
 }
```

@@

```{r heatmap.probs.bygene,cache=TRUE}
foo<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)x@z[,2]))
props<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)do.call(c,lapply(x@result@p,function(x)-diff(x[,2])))))
S<-sign(props)
S[S==0]<-1
colnames(S)<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])[,c("Patient.ID")]
stim<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])$Stim
rownames(S)<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)as.character(unique(pData(x)$Gene))))
heatmap.2((S*(foo))[apply(foo,1,function(x)sum(x>0.6)>=1),],col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="signed Posterior Probabilities\n(fit by gene, all stimulations)")
```

Figure 2 C, (fit of all samples simultaneously)

```{r eval=FALSE,echo=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/FluidigmCombinedFit.pdf")
foo2<-apply(foo,2,function(x)MIMOSA:::fdr(cbind(1-x,x)))
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
dev.off()
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
```

@@

```{r heatmap.props.bygene,cache=TRUE}
colnames(props)<-colnames(S)
rownames(props)<-rownames(S)
heatmap.2(S*sqrt(abs(props)),col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="Posterior Differences in Proportions\n(fit by gene all stimulations)")
```

@@

```{r heatmap.probs.bystim.bygene,cache=TRUE}
foo<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene,function(x)do.call(rbind,lapply(x,function(x)x@z[,2]))))
props<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene,function(x)do.call(rbind,lapply(x,function(x)do.call(c,lapply(x@result@p,function(x)-diff(x[,2])))))))
S<-sign(props)
colnames(S)<-do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])[,c("Patient.ID")])))
S[S==0]<-1
stim<-as.numeric(factor(do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])$Stim)))))
colnames(S)<-do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])$Patient.ID)))
rownames(S)<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene[[1]],function(x)as.character(unique(pData(x)$Gene))))
h<-heatmap.2((S*foo)[apply(foo,1,function(x)sum(x>.6)>=1),],col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],main="Signed Posterior Probabilities\n(fit by gene and stimulation).")
hmaporder<-h$colInd
geneinds<-apply(foo,1,function(x)sum(x>.6)>=1)
```

Figure 2 A

```{r eval=FALSE,echo=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Fluidigm_plot1.pdf")
foo2<-apply(foo,2,function(x)MIMOSA:::fdr(cbind(1-x,x)))
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
dev.off()
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
```

@@

```{r heatmap.props.bystim.bygene,cache=TRUE}
colnames(props)<-colnames(S)
rownames(props)<-rownames(S)
heatmap.2(sqrt(abs(props))*S,col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="Posterior differences in proportions\n(fit by gene and stimulations)")
```

@@

```{r cache=TRUE}
fisher<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x){
  p<-vector('numeric',nrow(x@result@n.stim))
  for(i in 1:nrow(x@result@n.stim)){
    p[i]<-fisher.test(as.matrix(rbind(x@result@n.stim[i,],x@result@n.unstim[i,])),alternative="two.sided")$p.value
  }
  p
}))
adj.fisher<-matrix(p.adjust(fisher,"fdr"),nrow=96)
S<-sign(lapply(fluidigm.fits.by.stim.allgenes,function(x)do.call(rbind,lapply(x,function(x)(prop.table(as.matrix(x@result@n.stim),1)-prop.table(as.matrix(x@result@n.unstim),1))[,2])))[[1]])
S[S==0]<-1
colnames(fisher)<-gsub("_B3GAT1_Treatment","",colnames(S))
rownames(fisher)<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)as.character(unique(pData(x)$Gene))))
stim<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])$Stim
heatmap.2(fisher*S,col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="-log10(p-values)\nfrom Fisher's Exact Test")
```

Figure 2 D

```{r eval=FALSE,echo=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/FluidigmFisher.pdf")
colnames(adj.fisher)<-colnames(fisher)
rownames(adj.fisher)<-rownames(fisher)
heatmap.2((((adj.fisher))*S)[apply(adj.fisher,1,function(x)sum(x<0.1)>1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
dev.off()
heatmap.2((((adj.fisher))*S)[apply(adj.fisher,1,function(x)sum(x<0.1)>1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
```

Figure 2C empirical proportions

```{r eval=FALSE,echo=FALSE}
emprop<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x){
  p<-vector('numeric',nrow(x@result@n.stim))
  p<-(prop.table(as.matrix((x@result@n.stim)),1)-prop.table(as.matrix(x@result@n.unstim),1))[,2]
}))
rownames(emprop)<-rownames(fisher)
colnames(emprop)<-colnames(fisher)
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/FluidigmEmpiricalProps.pdf")
heatmap.2(emprop[geneinds,hmaporder],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),Colv=NULL,denscol="black",tracecol="black",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)[hmaporder]])
dev.off()
heatmap.2(emprop[geneinds,hmaporder],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),Colv=NULL,denscol="black",tracecol="black",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)[hmaporder]])
```

@@

toplot<-data.frame(rbind(data.frame(CC$n.stim),data.frame(CC$n.unstim)),stim=c(rep("Stim",nrow(CC$n.stim)),rep("Unstim",nrow(CC$n.unstim))))
toplot$id<-rep(unlist(lapply(strsplit(as.character(subset(C,Stim.Condition%in%"Unstim")[,1]),":"),function(x)x[1]),use.names=FALSE),2)
colnames(toplot)[1:4]<-gsub("X","",paste(g1,g2,colnames(toplot)[1:4],sep=":"))
toplot<-melt(toplot,id=c("id","stim"))
p1<-ggplot(toplot)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous(na.value="gray")

DD<-CC
DD[[1]]<-cbind(CC[[1]][,1],rowSums(CC[[1]][,2:4]))
DD[[2]]<-cbind(CC[[2]][,1],rowSums(CC[[2]][,2:4]))
toplot2<-data.frame(rbind(data.frame(DD$n.stim),data.frame(DD$n.unstim)),stim=c(rep("Stim",nrow(DD$n.stim)),rep("Unstim",nrow(DD$n.unstim))))
toplot2$id<-rep(unlist(lapply(strsplit(as.character(subset(C,Stim.Condition%in%"Unstim")[,1]),":"),function(x)x[1]),use.names=FALSE),2)
colnames(toplot2)[1:2]<-gsub("X","",paste(g1,g2,colnames(toplot2)[1:2],sep=":"))
toplot2<-melt(toplot2,id=c("id","stim"))
p2<-ggplot(toplot2)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous(na.value="gray")

```{r fitmulti,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE,results='hide'}
flres<-.fitMCMC(data=CC,inits=MIMOSA:::MDMix(CC,initonly=TRUE),EXPRATE=1e-2,pXi=1,alternative="not equal")
flsum<-.fitMCMC(data=DD,inits=MIMOSA:::MDMix(DD,initonly=TRUE),EXPRATE=1e-2,pXi=1,alternative="not equal")
sapply(list.files(pattern=c("\\.dat")),unlink)
```

Figure 4

```{r eval=FALSE,echo=FALSE}
library(gridExtra)
missed<-!MIMOSA:::fdr(flres$z)<0.05
toplot$variable<-factor(toplot$variable,labels=c("CCR7-GZMK-","CCR7-GZMK+","CCR7+GZMK-","CCR7+GZMK+"))
toplot2$variable<-factor(toplot2$variable,labels=c("CCR7-GZMK-","CCR7+ or GZMK+"))
p1<-ggplot(toplot)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous("log2 count",na.value="gray",breaks=round(seq(0,6,l=10)))+theme(axis.text.y=element_text(color=c("black","red")[missed+1]))
p2<-ggplot(toplot2)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90,hjust=1))+facet_wrap(~stim)+scale_fill_continuous("log2 count",na.value="gray",breaks=round(seq(0,6,l=10)))+theme(axis.text.y=element_text(color=c("black","red")[2]))
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/revised_Fluidigm_Multivariate_CCR7_GZMK.pdf",width=6,height=4)
grid.arrange(p1,ncol=1)
dev.off()
grid.arrange(p1,ncol=1)
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/revised_Fluidigm_Marginal_CCR7_GZMK.pdf",width=6,height=4)
grid.arrange(p2,ncol=1)
dev.off()
grid.arrange(p2,ncol=1)

@@

ROCfun<-function(x,truth){
  o<-order(x,decreasing=FALSE)
  cbind(tpr=cumsum(truth[o])/sum(truth),fpr=cumsum(!truth[o])/sum(!truth))
}

fisher.p.adj<-lapply(fisher.p,function(x)p.adjust(x,"fdr"))
fisher.roc<-do.call(rbind,by(cbind(do.call(rbind,lapply(fisher.p.adj,function(i)ROCfun(i,gl(2,50)==2)))),rep(gl(100,1),10),function(x)colMeans(x)))
mimosa.adj<-lapply(1:ncol(mdsim),function(i)MIMOSA:::fdr(mdsim[,i]$z))
MIMOSA.roc<-do.call(rbind,by(do.call(rbind,lapply(mimosa.adj,function(i)ROCfun(i,gl(2,50)==2))),rep(gl(100,1),10),function(x)colMeans(x)))
foo<-data.frame(as.data.frame(rbind(fisher.roc,MIMOSA.roc)),Method=gl(2,100,labels=c("Fisher","MIMOSA")))
sz<-element_text(size=18)
p1<-ggplot(foo)+geom_line(aes(x=fpr,y=tpr,col=Method))+scale_y_continuous("TPR")+scale_x_continuous("FPR")+theme_bw()+theme(axis.title.x=sz,axis.title.y=sz,axis.text.x=sz,axis.text.y=sz)
#FDR
mim.fdr<-as.data.frame(cbind(do.call(rbind,lapply(mimosa.adj,function(x)fdrCurves(x,gl(2,50)==2))),rep=gl(100,1)))
ddply(mim.fdr,.(rep),summarize,true.fdr.hat=mean(true.fdr),fdr.hat=mean(fdr))
fisher.fdr<-as.data.frame(cbind(do.call(rbind,lapply(fisher.p.adj,function(x)fdrCurves(x,gl(2,50)==2))),rep=gl(100,1)))
plot(ddply(mim.fdr,.(rep),summarize,true.fdr.hat=mean(true.fdr),fdr.hat=mean(fdr)))
```

Figure 5

```{r eval=FALSE,echo=FALSE}
```

@@

Figure S3
